---
title: "master_clean_shape"
author: "Fanmei Xia"
date: "2/12/2021"
output:
  pdf_document: default
  html_document:
    df_print: paged
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{R, warning = F, error = F}
library(tidyverse)
library(tidylog)
library(sf)
library(tmap)
 
path <- "/Users/xiafm/Documents/GitHub/NJ-opioidenv/data_final/"
master_clean <- read.csv(paste0(path, "master_clean.csv"), colClasses = c("SSN" = "character"))

mun <- st_read("data_in_progress/mun_boundaries.geojson") %>%
  select(SSN) 

master_shape <- left_join(master_clean, mun) %>%
  st_as_sf()

```

```{r, warning = F, error = F}
col_names <- colnames(master_shape)
col_names <- col_names[c(5, 10:21, 23:29, 39:41)]
print(col_names)

make_map <- function(col_name) {
  tm_shape(master_shape) + tm_fill(col = col_name, style = "jenks") + tm_borders(alpha = .15)
}
maps <- lapply(col_names, make_map)
```

## Commercial Environment:
Bars/Liquor store per capita and per area: 
8. count_bars
9. count_ls
10. bars_per_sqft
11. bars_per_pop
12. ls_per_sqft
13. ls_per_pop
14. bars_ls_per_pop

Comments: 
1. very limited data points in selected areas;
2. per pop seems to be a better variable?

```{r, warning=FALSE, error=FALSE, message=FALSE}
par(mfrow=c(1,2))
maps[8] #8. count_bars
maps[9] #9. count_ls

par(mfrow=c(3,2))
maps[10] #10. bars_per_sqft
maps[11] #11. bars_per_pop
maps[12] #12. ls_per_sqft
maps[13] #13. ls_per_pop
maps[14] #14. bars_ls_per_pop
```

##Residential Environment:

Comment: 
1. school measures don't have enough data
2. public transit most availabel in central area
3. relatively consistent patterns that the souther side of NJ has a higher rate of mobile home, vacancy rate, 
Takeaways: Again relatively consistent patterns. Appears that sales volume was relatively consistent but went down in rural western side. 
```{r, warning = F, error = F}
maps[4] #4	num_schools
maps[5] # 5	schools_per_pop
maps[15] #15 multiunit_struct"  
maps[16] #16 no_vehicle 
maps[17] #17 public_transit
maps[18] #18 mobile_home
maps[19] #19 total_units 
maps[20] #20 vacant_units
maps[21] #21 occupancy_rate
maps[22] #22 vacancy_rate
maps[23] #23 mobile_home_rate
```

##Physical Environ:

Comment:
1. we may need better measures?
```{r, warning = F, error = F}
maps[1] #1. 	pct_h_den_res
maps[2] #2. med_walk_score
maps[3] #3. mean_walk_scre
maps[6] #6. bike_path_ft_p_mile"
maps[7] #7. "bikes_ft_p_mile"

```

