---
title: "Visualizing_LU_and_Business"
author: "Gabe Morrison"
date: "10/18/2020"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r, echo = F}
library(tidyverse)
library(tidylog)
library(sf)
library(tmap)

#Sorry this is not reproducible. Will need to change to your directory but keep /NJ-opioidenv/data_raw
setwd("~/Documents/U_Chicago_Year_4/NJ-opioidenv/data_in_progress")
pe <- st_read('physical_environment_2015.geojson')
```

```{r}
#Sorry this is not reproducible. Rmarkdown throws me an error otherwise and this is a workaround. 
setwd("~/Documents/U_Chicago_Year_4/NJ-opioidenv/data_in_progress")
pe <- st_read('physical_environment_2015.shp')
pe2 <-st_read('physical_environment_2015.gpkg')
```

##Task 1: Visualize Land Use to compare urban vs total area measures
Specifically, try to answer question: which measure should we use?

####Commercial:
```{r}
commercial_urban <- tm_shape(pe) + tm_borders() + tm_fill(col = "pct_com_urb") + tm_layout(legend.outside = TRUE)
commercial_total <- tm_shape(pe) + tm_borders() + tm_fill(col = "pct_com_tot") + tm_layout(legend.outside = TRUE)
commercial_urban
commercial_total
```

####Industrial:
```{r}
industrial_urban <- tm_shape(pe) + tm_borders() + tm_fill(col = "pct_ind_urb") + tm_layout(legend.outside = TRUE)
industrial_total <- tm_shape(pe) + tm_borders() + tm_fill(col = "pct_ind_tot") + tm_layout(legend.outside = TRUE)
industrial_urban
industrial_total
```


####Residential:
```{r}
residential_urban <- tm_shape(pe) + tm_borders() + tm_fill(col = "pct_res_urb") + tm_layout(legend.outside = TRUE)
residential_total <- tm_shape(pe) + tm_borders() + tm_fill(col = "pct_res_tot") + tm_layout(legend.outside = TRUE)
residential_urban
residential_total
```


####Takeaway:
Huge left skew in commercial and residential of low values. This makes sense; in rural areas there are not very many major commercial/industrial areas. Visualizing the residential land, unsurprisingly in much of the rural areas, there were substantially higher percents of residential land use/urban. This was not as high for residential/total.

(1) Considering total area considers nature (ie places where people just don't live)
(2) Less-skewed distributions, especially for commerical and industrial 


###Task 2: Commerical Districts:
```{r}
pe <- pe %>%
  mutate(mun_type_num = case_when(
    mun_type == "Township" ~ 1,
    mun_type == "Borough" ~ 2, 
    mun_type == "City" ~ 3
  ))
nominal_density <- tm_shape(pe) + tm_fill(col = "mun_type_num")
nominal_density
```

Note: This is not a great measure because these are forms of government and do not reflect scale in any sense. Therefore, not good measure of central business. https://www.njlm.org/644/Forms-of-Municipal-Government---New-Jers

