---
title: "Opioid Project Meeting Jan 21, 2020"
author: "Gabe Morrison"
date: "1/18/2021"
output:
  html_document:
    df_print: paged
  pdf_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

### Concerns about Data:
**Franchise vs individual establishment:** The data is listed by individual establishment. For example, in the first 400 rows, there are 7 McDonalds (in PR). 

**Square footage:** I double checked and didn't find this. There is an Office Size Code column but it is empty. 

### Other Notes:
Data also contains the percent of businesses in each municipality that are in each 2-digit NAICS classification. I did not visualize those for eac

```{R, warning = F, error = F}
library(tidyverse)
library(tidylog)
library(sf)
library(tmap)
 
bus_data <- read.csv("data_in_progress/nj_bus_mun.csv",
                     colClasses = c("SSN" = "character")) 

mun <- st_read("data_in_progress/mun_boundaries.geojson") %>%
  select(SSN) 

bus_data <- left_join(bus_data, mun) %>%
  st_as_sf
```

```{r, warning = F, error = F}
col_names <- colnames(bus_data)
col_names <- col_names[-(1:2)]
col_names <- col_names[-97]

make_map <- function(col_name) {
  tm_shape(bus_data) + tm_fill(col = col_name, style = "jenks") + tm_borders(alpha = .15)
}
maps <- lapply(col_names, make_map)
```


## Emp Size 18, 14, dif:

Takeaways: Generally small employment size near NYC; mean size increased slightly in rural areas, not a ton in urban NE and E. 
```{r, error =F}
maps[1]
maps[2]
maps[33]
maps[34]
maps[65]
maps[66]
```


##Sales Volume:

Takeaways: Again relatively consistent patterns. Appears that sales volume was relatively consistent but went down in rural western side. 
```{r, warning = F, error = F}
maps[3]
maps[4]
maps[35]
maps[36]
maps[67]
maps[68]
```


##Simpson Index Number of employees:

Relatively high diversity across the state. Not a lot of variability in the measures :(, so may not be great for modelling.
```{r, warning = F, error = F}
maps[5]
maps[37]
maps[69]
```

##Business Counts:

Takeaway: Consistently more businesses where there are more people, even as those municipalities are smaller. These data don't change a ton 2014-2018 (light yellow is close to no change). 
```{r, warning = F, error = F}
maps[32]
maps[64]
maps[96]
```


## Using NAICS to identify urban:
Not terribly successful; may need raw counts rather than percents

### NAICS 54- Professional, Scientific, and Technical Services:

```{r, warning = F, error = F}
maps[18]
maps[52]
maps[84]
```



#NAICS 55- Management of Companies and Enterprises:

```{r, warning = F, error = F}
maps[21]
maps[53]
maps[85]
```


#NAICS 71- Arts, Entertainment, and Recreation:

```{r}
maps[23]
maps[57]
maps[89]
```